---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# airbnbtools

<!-- badges: start -->

[![codecov](https://codecov.io/gh/DSCI-310-2025/airbnbtools/graph/badge.svg?token=M05R361SU6)](https://codecov.io/gh/DSCI-310-2025/airbnbtools)
[![R-CMD-check](https://github.com/DSCI-310-2025/airbnbtools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DSCI-310-2025/airbnbtools/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](LICENSE)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/)

<!-- badges: end -->


The goal of **airbnbtools** is to streamline the process of working with Airbnb datasets, offering tools for downloading, transforming, and modeling data. This package provides several functions to automate tasks such as data acquisition, data cleaning, and model training and evaluation, helping to make the data analysis workflow more efficient.

You can find the documentation for this package at the following URL:

[airbnbtools Documentation](https://dsci-310-2025.github.io/airbnbtools/)

## Installation

You can install the development version of **airbnbtools** from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("DSCI-310-2025/airbnbtools")
```

## Usage

The `airbnbtools` package provides several useful functions for data manipulation and analysis. Here are some common tasks:

### Convert selected columns to factor type

The function `convert_columns_to_factor()` converts specific columns of a data frame to factors.

```{r}
df <- data.frame(
  room_type = c("Private room", "Entire home/apt"),
  host_is_superhost = c("t", "f"),
  city = c("Madrid", "Barcelona"),
  weekdays = c(TRUE, FALSE),
  price = c(100, 150)
)
convert_columns_to_factor(df)
```

### Download and combine Airbnb data

Use `download_combine_data()` to download data from multiple sources and combine them into a single data frame.

```{r eval=FALSE}
urls <- list(
  "Vancouver_weekdays" = "https://example.com/vancouver_weekdays.csv",
  "Toronto_weekends" = "https://example.com/toronto_weekends.csv"
)
combined_df <- download_combine_data(urls)
head(combined_df)
```

### Create a grid of histograms

The function `create_histogram_grid()` generates and saves a grid of histograms for numeric columns in a data frame.

```{r}
df <- data.frame(
  price = c(100, 150, 200, 250),
  size = c(20, 25, 30, 35),
  rating = c(4.5, 4.0, 5.0, 3.5)
)
out_path <- tempfile(fileext = ".png")
create_histogram_grid(df, vars = c("price", "size", "rating"), output_path = out_path)
```

### Evaluate a regression model

Use `evaluate_model()` to compute evaluation metrics (RMSE, MAE, R-squared) for a regression model.

```{r}
df <- data.frame(x = 1:10, y = 2 * (1:10) + rnorm(10))
model <- lm(y ~ x, data = df)
result <- evaluate_model(model, df, response_col = "y")
print(result)
```

### Split data into training and testing sets

The `split_train_test()` function splits a dataset into training and testing sets using stratified sampling.

```{r}
df <- data.frame(x = rnorm(100), y = rnorm(100))
split_data <- split_train_test(df, response_col = "y", train_ratio = 0.75)
str(split_data$train)
str(split_data$test)
```

### Train a regression model using caret

Use `train_model()` to train a regression model using the `caret` package.

```{r}
df <- data.frame(
  price = rnorm(100, mean = 200, sd = 50),
  size = rnorm(100, mean = 80, sd = 10),
  rating = rnorm(100, mean = 4, sd = 0.5)
)
model <- train_model(df)
summary(model)
```

### Validate Airbnb data

The `validate_airbnb_data()` function checks and converts specified columns of a data frame to factors.

```{r eval=FALSE}
df <- data.frame(
  room_type = c("Entire home/apt", "Private room"),
  host_is_superhost = c("t", "f"),
  city = c("Berlin", "Paris"),
  weekdays = c("Monday", "Tuesday")
)
validate_airbnb_data(df)
```
